---
title: "Explorative Data Analysis"
author: "Enrico V"
date: "April 2018"
output: html_document
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE
)
```


# Foreword  
Goal of EDA (exploratory Data Analysis) is to get a general idea of the data.
One of the key principles taught in the course, probably the key principle is **"start/lead with the question"**.
The question is clearly defined by the objective of the capstone project: next
word predictions. We are also given some requirements, not of interest for non
specialists, that make very clear that information about volume of data (ex. size
and length of text, number of different workds) might be important.
Here I try to follow that principles, so this analysis might be slightly
different from examples of text  EDA available in courses books and internet as they have a different focus("question") than ours. For instance:  a basic didactic example is the "question" to predict whether an email is spam or not. 
For that questions "neuter" and/or common words are not  useful (ex. "stopwords") while less common words could be useful and meaningful.
For our "question"/goal it seems that we cannot ignore stopwords. 
An exploratory analysis is not a final place to draw conclusion but an initial place to understand the data, so  rather than discard "stopwords" up front as other analysis probably would, in our EDA it seems important to get information about stopwords, information useful at later stages to decide what to do about them,




```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      echo=FALSE, warning=FALSE, message=FALSE)
```

```{r load_libraries, echo=FALSE}
require(quanteda)
require(readtext)
source("01_globals.R")
```



```{r load_files, echo=TRUE}
#fnames_pattern <- file.path(data_dir_corpus,"*subset*")

#blogs <- readtext(fnames_pattern
# ,docvarsfrom = "filenames", dvsep = "-"
# ,docvarnames = c("lang","title","subset","nlines_kept","nlines_read"))

#myCorpus <- corpus(blogs)
#print(summary(myCorpus))
```



